---
title: "Spatial Distribution of Traumatic Brain Injury Contusions<br>and Post-Traumatic Epilepsy"
author:
  - name: Daniel Brennan PhD
    affil: 1
  - name: Andrea L.C. Schneider MD PhD
    affil: 1
  - name: James J. Gugger MD PharmD
    affil: 2

affiliation:
  - num: 1
    address: University of Pennsylvania Department of Neurology, Philadelphia PA, United States
  - num: 2
    address: University of Rochester Department of Neurology, Rochester NY, United States

affiliation_textsize: 25
column_numbers: 3

primary_colour: "#011F5B"
secondary_colour: "#FFFFFF"
main_textcol: "#FFFFFF"
body_textcol: "#000000"
affiliation_textcol: "#FFFFFF"
title_height_ratio: 0.175

output:
  posterdown::posterdown_html:
    self_contained: false
bibliography: packages.bib
---

```{r setup, include=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(readr))
suppressMessages(library(patchwork))
suppressMessages(library(knitr))
suppressMessages(library(png))

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%"
)

poster_dir <- dirname(normalizePath(knitr::current_input(), mustWork = TRUE))

pte_cols <- c("non-PTE" = "#1f77b4", "PTE" = "#d62728")
panel_title_map <- c(
  "core + edema" = "Core + Edema",
  "core" = "Core",
  "edema" = "Edema",
  "disconnection mapping" = "White Matter Disconnection"
)

glass_stat_file_map <- c(
  "core + edema" = "glass_stat_both_5_epi.png",
  "core" = "glass_stat_core_5_epi.png",
  "edema" = "glass_stat_edema_5_epi.png",
  "disconnection mapping" = "glass_stat_disconnectome_5_epi.png"
)

plot_cosine_panel <- function(df, panel_title, show_x_axis = TRUE) {
  p <- ggplot(df, aes(x = pte_status, y = residualized_cosine, color = pte_status)) +
    geom_boxplot(outlier.shape = NA, width = 0.55, alpha = 0.20, linewidth = 1.3) +
    geom_jitter(width = 0.10, alpha = 0.8, size = 3.0) +
    scale_color_manual(values = pte_cols, drop = FALSE) +
    labs(
      title = panel_title,
      x = "PTE Status",
      y = "Residualized Cosine Similarity"
    ) +
    theme_minimal(base_size = 22) +
    theme(
      legend.position = "none",
      plot.title = element_text(size = 25, face = "bold", hjust = 0.5),
      axis.title.x = element_text(size = 19),
      axis.title.y = element_text(size = 19),
      axis.text = element_text(size = 17),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.margin = margin(12, 5, 12, 5)
    )

  if (!show_x_axis) {
    p <- p +
      theme(
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
      )
  }

  p
}

missing_cosine_panel <- function(panel_title, note) {
  ggplot() +
    annotate("text", x = 0.5, y = 0.62, label = panel_title, size = 7, fontface = "bold") +
    annotate("text", x = 0.5, y = 0.40, label = note, size = 5) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE) +
    theme_void()
}

plot_glass_panel <- function(image_path) {
  img <- png::readPNG(image_path)

  ggplot() +
    annotation_custom(
      grid::rasterGrob(img, interpolate = TRUE),
      xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = NA),
      plot.margin = margin(12, 12, 12, 5)
    )
}

missing_glass_panel <- function(panel_title, note) {
  ggplot() +
    annotate("text", x = 0.5, y = 0.58, label = panel_title, size = 7, fontface = "bold") +
    annotate("text", x = 0.5, y = 0.40, label = note, size = 5) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "#cccccc", linewidth = 0.4),
      plot.margin = margin(12, 12, 12, 5)
    )
}
```

<style>
.poster_title h3 {
  margin-top: 0.28em !important;
}

.poster_title h5 {
  margin-top: 0.12em !important;
}

.poster_body ul {
  margin-top: 0.35em;
  margin-bottom: 0.45em;
}

.poster_body ul li {
  margin-bottom: 0.28em;
  line-height: 1.18;
}

.poster_body ul li:last-child {
  margin-bottom: 0;
}
</style>

# Background

* Focal traumatic brain injury (TBI) and contusions are associated with post-traumatic epilepsy (PTE)
* Observed differences confounded by increased injury severity
* Unclear whether focal injury related beyond injury severity, which contusion characteristics (core/edema) associated

Objective: To characterize the spatial distribution of TBI contusions in PTE

# Methods

* Transforming Research and Clinical Knowledge in TBI Epileptogenesis Project (TRACK-TBI EPI) data
* Manual contusions segmentation on 2-week T1/FLAIR MRI; masks warped to MNI space
* Disconnection mapping: extend contusion via intersection of normative white-matter tract atlas
* LESYMAP: multivariate contusion map analysis
* Cosine-similarity of stat-map & contusions compared with covariates: age, sex, GCS, total-contusion-volume

# Results

```{r demographics-table, results='asis'}
demo_table_path <- file.path(poster_dir, "table1_demographics_poster.csv")

if (!file.exists(demo_table_path)) {
  cat("<p><em>Run my_lesymap.Rmd to export the demographics table.</em></p>")
} else {
  demo_table <- readr::read_csv(demo_table_path, show_col_types = FALSE)

  cat("<div style='height:0.55em;'></div>")
  cat("<div style='font-size:20pt; font-weight:700; margin-top:0.20em; margin-bottom:0.42em;'>Demographics</div>")

  cat("
<style>
table.apa-table {
  width: 100%;
  border-collapse: collapse;
  border-top: 2px solid #000000;
  border-bottom: 2px solid #000000;
  font-size: 20pt;
  line-height: 1.15;
  margin-top: 0.15em;
}
table.apa-table th,
table.apa-table td {
  padding: 0.36em 0.48em;
  border-left: none;
  border-right: none;
}
table.apa-table thead th {
  border-bottom: 1.5px solid #000000;
  font-weight: 700;
}
table.apa-table tbody tr:last-child td {
  border-bottom: none;
}
</style>
")

  apa_table <- knitr::kable(
    demo_table,
    format = "html",
    escape = FALSE,
    align = c("l", "c", "c", "c"),
    table.attr = 'class="apa-table"'
  )
  cat(apa_table)
}
```

```{r lesymap-results-table, results='asis'}
results_table <- data.frame(
  `Contusion Characteristic` = c(
    "Core + Edema",
    "Core",
    "Edema",
    "Disconnection"
  ),
  PTE = c("14", "14", "10", "13"),
  `non-PTE` = c("121", "121", "90", "129"),
  `LESYMAP p-value` = c("0.004", "0.069", "0.018", "0.006"),
  `Median Difference<br>Cosine Similarity [95% CI]` = c(
    "0.013 [0.001, 0.031]",
    "0.046 [0.012, 0.096]",
    "0.026 [-0.008, 0.061]",
    "0.096 [0.038, 0.148]"
  ),
  `p-value` = c("0.036", "0.010", "0.131", "0.003"),
  check.names = FALSE
)

cat("<div style='height:1.05em;'></div>")
cat("<div style='font-size:20pt; font-weight:700; margin-top:0.18em; margin-bottom:0.42em;'>Contusion Mapping Results</div>")

results_kable <- knitr::kable(
  results_table,
  format = "html",
  escape = FALSE,
  align = c("l", "c", "c", "c", "c", "c"),
  table.attr = 'class=\"apa-table\"'
)
cat(results_kable)
```

# Conclusions

* Contusions burden increased in PTE beyond injury severity - need for external validation
* White matter disconnected by contusions represents strong candidate biomarker
* Contusion-disconnected white matter may carry PTE relevant signal beyond focal injury burden

# Results {.mybreak .span}

```{r cosine-panels, fig.width=22, fig.height=20}
cosine_data_path <- file.path(poster_dir, "poster_cosine_panels.csv")
panel_levels <- c("core + edema", "core", "edema", "disconnection mapping")

if (!file.exists(cosine_data_path)) {
  panel_plots <- lapply(
    panel_levels,
    function(lbl) missing_cosine_panel(lbl, "Run my_lesymap.Rmd to export poster cosine data.")
  )
} else {
  cosine_panel_df <- readr::read_csv(cosine_data_path, show_col_types = FALSE) |>
    mutate(
      pte_status = factor(pte_status, levels = c("non-PTE", "PTE")),
      panel_label = factor(panel_label, levels = panel_levels)
    )

  panel_plots <- lapply(seq_along(panel_levels), function(i) {
    panel_name <- panel_levels[[i]]
    panel_df <- cosine_panel_df |>
      filter(panel_label == panel_name, !is.na(residualized_cosine), !is.na(pte_status))

    if (nrow(panel_df) == 0) {
      missing_cosine_panel(panel_title_map[[panel_name]], "No exported rows for this panel.")
    } else {
      plot_cosine_panel(panel_df, panel_title_map[[panel_name]], show_x_axis = i == length(panel_levels))
    }
  })
}

glass_panels <- lapply(panel_levels, function(panel_name) {
  image_path <- file.path(poster_dir, glass_stat_file_map[[panel_name]])

  if (!file.exists(image_path)) {
    missing_glass_panel(panel_title_map[[panel_name]], "Add corresponding glass-stat PNG to the poster folder.")
  } else {
    plot_glass_panel(image_path)
  }
})

paired_rows <- lapply(seq_along(panel_levels), function(i) {
  glass_panels[[i]] | panel_plots[[i]]
})

wrap_plots(paired_rows, ncol = 1) +
  plot_layout(widths = c(2, 1))
```
